---
title: "BeeREX implementation in R"
author: "Andrew Kniss"
output: html_document
---

From the [EPA description](https://www.epa.gov/pesticide-science-and-assessing-pesticide-risks/models-pesticide-risk-assessment): 

> "The Bee-REX model is a screening level tool that is intended for use in a Tier I risk assessment to assess exposures of bees to pesticides and to calculate risk quotients. This model is individual-based, and is not intended to assess exposures and effects at the colony-level (i.e., for honey bees)."

```{r setup, include=FALSE}
library(tidyverse)
### Daily consumption of food for different bee stages (Table 4)
### to be used in dietary dos and RQ calculations
diet.table <- data.frame(
  jelly.mgd = c(1.9, 9.4, 19, 0, 0, 0, 1.9, 9.4, 23, 141, 0, 0, 0, 0, 0, 0, 0, 525),
  nectar.mgd = c(0, 0, 0, 60, 120, 130, 0, 0, 0, 0, 60, 140, 60, 43.5, 292, 29, 235, 0),
  pollen.mgd = c(0, 0, 0, 1.8, 3.6, 3.6, 0, 0, 0, 0, 6.65,
                 9.6, 1.7, 0.041, 0.041, 2, 0.0002, 0),
  stage = c(rep("larval", 10), rep("adult", 8)))

## EEC calculation functions in ug ai/mg (Table 3)      ##
## multiply by 1,000 to get EEC in mg ai/kg             ##
eec.foliar <- function(rate.kgha) {
  rate.kgha * 98 / 1000
}
eec.soil <- function(rate.kgha, log.Kow = 5, Koc = 30) {
  ((10^(0.95*log.Kow-2.05)+0.82) * 
     (-0.0648*(log.Kow^2)+0.2431*log.Kow+0.5822) *
     (1.5/(0.2+1.5*Koc*.01)) * (0.45*rate.kgha)) / 1000
}
eec.seed <- function(rate.kgha) {
  1
}
### Total dose function in ug/bee
### Produces a list of values equal to total dose column in Table 4
total.dose <- function(EEC.ugmg,
                       jelly.mgd = diet.table$jelly.mgd,
                       nectar.mgd = diet.table$nectar.mgd,
                       pollen.mgd = diet.table$pollen.mgd) {
  ug.bee <- (EEC.ugmg * pollen.mgd + 
               EEC.ugmg * nectar.mgd + 
               EEC.ugmg/100 * jelly.mgd)
  #names(ug.bee) <- .$pesticide
  return(ug.bee)
}

### Input file with necessary data ##
### (contains data from Tables 1 & 2):     ##
pestDF <- read.csv("BeeRexInputValues-test1.csv") %>%
  mutate(rate.kgha = 1,
         appType = "soil")
glimpse(pestDF)
```

This implementation allows rapid calculation of acute and chronic risk quotients (RQ) for many pesticides, as long as the following information is assembled for each pesticide: 

* application rate (kg/ha)
* application type (foliar, soil, or seed)
* log K~ow~ (default to 5 if not provided)
* K~oc~ (defaults to 30 if not provided)
* adult honeybee contact LD~50~ 
* adult honeybee oral LD~50~
* adult honeybee chr2onic NOAEL
* larval honeybee oral LD~50~
* larval honeybee chronic NOAEL

An example data set with two pesticides has been included for QA/QC purposes.
```{r echo=FALSE}
knitr::kable(pestDF)
```

Based on those input values, the tool will estimate dietary exposure (in ub per bee) for these two pesticides for a range of honeybee stages:

```{r, echo=FALSE}
pestDF.2 <- pestDF %>%
  mutate(eec = case_when(
    appType == "foliar" ~ eec.foliar(rate.kgha),
    appType == "soil" ~ eec.soil(rate.kgha, log.Kow, Koc),
    appType == "seed" ~ eec.seed(rate.kgha))) %>%
    mutate(tab4dose = {map(.x = .$eec, .f = total.dose)}) 
```

The dietary exposure values are then divided by the appropriate toxicological endpoint (LD~50~ or NOAEL) to obtain acute and chronic risk quotients. For each stage, the highest RQ value is then selected to be used in subsequent modeling. 

```{r, echo=FALSE, message=FALSE, warning = FALSE}
pestDF.3 <- pestDF.2 %>%
    unnest(cols = c(tab4dose)) %>%
    mutate(stage = rep(diet.table$stage, length(unique(commonName))),
           RQ.acute = case_when(
               stage == "larval" ~ tab4dose / LD50.larva.oral,
               stage == "adult" ~ tab4dose / LD50.adult.oral),
           RQ.chronic = case_when(
               stage == "larval" ~ tab4dose / NOAEL.larva.oral,
               stage == "adult" ~ tab4dose / NOAEL.adult.oral)) %>%
    group_by(commonName, CAS, stage) %>%
    summarize(max.acute.RQ = max(RQ.acute, na.rm=TRUE),
              max.chronic.RQ = max(RQ.chronic, na.rm=TRUE)) %>%
    pivot_wider(id_cols = c(commonName, CAS),
                names_from = stage,
                values_from = c(max.acute.RQ, max.chronic.RQ))
finalDF <- left_join(pestDF.2, pestDF.3) %>%
  mutate(max.contact.RQ_adult = 
         case_when(appType == "foliar" ~ rate.kgha * 2.4 / 
                     LD50.adult.contact)) %>%
  select(CAS, commonName, rate.kgha, appType,
         max.acute.RQ_adult, max.acute.RQ_larval,
         max.chronic.RQ_adult, max.chronic.RQ_larval,
         max.contact.RQ_adult) 

knitr::kable(finalDF)
```
